---
title: "School buildings in my earthquake field research"
author: "Behtash JavidSharifi"
date: "2026-09-23"
categories: [Earthquake Engineering, RECoVER, R]
format:
  html:
    code-fold: false
execute:
  echo: true
  cache: false
  freeze: false
---

## Why describe the building sample?

My PhD research examines relationships between earthquake ground motion,
structural response, and observed building damage.
As part of this work, RECoVER brings together field evidence and numerical
analysis.

This post describes the building sample recorded in one snapshot of my
field ambient-vibration review workbook.
Before examining vibration characteristics, I want to understand which
buildings and damage categories the dataset represents.

## Data source and scope

The data come from the `CASE_REVIEW` sheet of my
`RECoVER_FieldAV_AdjudicationWorkbook_AR38D.xlsx` research workbook.

The [building-data CSV](field-av-buildings.csv) contains the selected
columns used here.
It retains the workbook's building identifiers, reported storey counts,
damage labels, and evidence routes.

This is a descriptive analysis of a research sample.
It does not establish that building height caused a particular damage level.
The workbook's modal selections remain a separate review task.

## Load the packages

I suppress package startup messages to keep the page readable.
The analysis code and its results remain visible.

```{r}
#| label: setup
#| message: false

library(readr)
library(dplyr)
library(ggplot2)
library(knitr)
```

## Read and inspect the building records

Each row represents one building case.
A station can be associated with more than one building, so I use
`CaseID` to identify individual records.

```{r}
#| label: inspect-buildings

buildings <- read_csv(
  "field-av-buildings.csv",
  show_col_types = FALSE
)

buildings |>
  select(CaseID, StoreysReported, FieldDamageLabel) |>
  kable(
    col.names = c("Building case", "Reported storeys", "Recorded damage")
  )
```

## Check completeness and identifiers

Before calculating summaries, I check for missing values and repeated
building identifiers.

```{r}
#| label: check-building-data

data_checks <- buildings |>
  summarise(
    building_records = n(),
    unique_buildings = n_distinct(CaseID),
    missing_storeys = sum(is.na(StoreysReported)),
    missing_damage_labels = sum(is.na(FieldDamageLabel))
  )

kable(data_checks)
```

## Summarize the recorded damage categories

The table shows how many buildings fall in each recorded damage category,
together with their mean and range of reported storey counts.

```{r}
#| label: summarize-buildings

damage_summary <- buildings |>
  group_by(FieldDamageLabel) |>
  summarise(
    buildings = n(),
    mean_storeys = mean(StoreysReported),
    minimum_storeys = min(StoreysReported),
    maximum_storeys = max(StoreysReported),
    .groups = "drop"
  )

kable(
  damage_summary,
  digits = 2,
  col.names = c(
    "Recorded damage", "Buildings", "Mean storeys",
    "Minimum storeys", "Maximum storeys"
  )
)
```

## Compare the building sample visually

Point size represents the number of buildings sharing the same damage
label and reported storey count.

Fractional storey counts are preserved as reported in the workbook.
They have not been rounded or reinterpreted.

```{r}
#| label: fig-building-storeys
#| fig-cap: "Reported storey counts by recorded damage category. Larger points indicate more buildings at the same position."
#| fig-width: 8
#| fig-height: 4.5

ggplot(
  buildings,
  aes(x = StoreysReported, y = FieldDamageLabel)
) +
  geom_count(colour = "#2563EB") +
  scale_size_area(max_size = 9, breaks = c(1, 2)) +
  labs(
    x = "Reported number of storeys",
    y = "Recorded damage category",
    size = "Buildings"
  ) +
  theme_minimal(base_size = 12)
```

## What does this snapshot show?

The sample includes buildings with several recorded damage levels.
Moderate and low damage are the most common labels in this snapshot,
while severe damage is represented by only one building.

The categories also overlap in reported storey count.
These descriptive results do not isolate the effects of structural
design, construction quality, or earthquake exposure.

For my research, this overview helps document the sample before moving
to the separate tasks of reviewing vibration evidence and comparing
structural models.