---
title: "From R to Python and back: reviewing vibration-frequency recurrence"
author: "Behtash JavidSharifi"
date: "2026-09-23"
categories: [Earthquake Engineering, RECoVER, R, Python]
engine: knitr
format:
  html:
    code-fold: false
execute:
  echo: true
  cache: false
  freeze: false
---

## Why combine R and Python?

My doctoral research uses ambient-vibration evidence to help characterize
buildings affected by earthquakes. This post demonstrates a workflow in
which R summarizes candidate frequency families, Python calculates their
relative frequencies, and R presents the returned results.

The objects pass directly between the two languages through reticulate.
No intermediate CSV is written or reloaded during that exchange.

## Data source and scope

The data come from the `FAMILIES` sheet of my
`RECoVER_FieldAV_AdjudicationWorkbook_AR38D.xlsx` research workbook.

I reuse the
[frequency-family CSV](../ambient-vibration-repeatability/field-av-frequency-families.csv)
introduced in my
[Python analysis](../ambient-vibration-repeatability/index.qmd).

Each row describes a candidate frequency family within a building case.
These families are not automatically accepted structural modes.

This preliminary research sample is shared to demonstrate a reproducible
analysis workflow for DSCI 521 Milestone 3 and my research portfolio.
It represents a selected workbook snapshot, rather than the complete
or final research dataset. The results should not be interpreted as
final thesis findings.

## Connect the project environments

I select the Python interpreter inside this project's virtual environment
before initializing Python. Package startup messages are suppressed to
keep the page readable; code and analytical results remain visible.

```{r}
#| label: setup
#| message: false

project_root <- Sys.getenv("QUARTO_PROJECT_DIR")
stopifnot(nzchar(project_root))

python_path <- file.path(
  project_root,
  ".venv",
  if (.Platform$OS.type == "windows") {
    "Scripts/python.exe"
  } else {
    "bin/python"
  }
)

Sys.setenv(
  RETICULATE_PYTHON = normalizePath(
    python_path, winslash = "/", mustWork = TRUE
  )
)

library(reticulate)
library(readr)
library(dplyr)
library(knitr)
```

## Summarize the records in R

I check the recording-window counts and count the families represented
in each recurrence category.

```{r}
#| label: summarize-in-r

families <- read_csv(
  file.path(
    project_root,
    "posts",
    "ambient-vibration-repeatability",
    "field-av-frequency-families.csv"
  ),
  show_col_types = FALSE
)

stopifnot(
  nrow(families) > 0,
  !anyNA(families$WindowsRepresented),
  all(families$WindowsRepresented %in% 1:3)
)

recurrence_r <- families |>
  count(WindowsRepresented, name = "Families") |>
  arrange(WindowsRepresented) |>
  as.data.frame()

kable(
  recurrence_r,
  col.names = c("Windows represented", "Candidate families")
)
```

## Pass the R summary to Python

Python accesses the R object through `r.recurrence_r`. It calculates
each category's percentage of all candidate families and the percentage
represented in at least two windows.

```{python}
#| label: calculate-in-python

import pandas as pd

recurrence_py = r.recurrence_r.copy()
assert isinstance(recurrence_py, pd.DataFrame)

recurrence_py["Percentage"] = (
    100 * recurrence_py["Families"]
    / recurrence_py["Families"].sum()
)

repeated_percentage = float(
    recurrence_py.loc[
        recurrence_py["WindowsRepresented"] >= 2,
        "Percentage"
    ].sum()
)

print(recurrence_py.to_string(index=False))
```

## Bring the Python result back to R

R accesses the Python objects through `py$`. The percentage column
below was computed in Python, rather than recalculated in R.

```{r}
#| label: present-python-results-in-r

returned_summary <- py$recurrence_py

stopifnot(
  isTRUE(all.equal(
    sum(returned_summary$Percentage),
    100,
    tolerance = 1e-8
  ))
)

kable(
  returned_summary,
  digits = 1,
  col.names = c(
    "Windows represented",
    "Candidate families",
    "Percentage of families"
  ),
  caption = "Counts computed in R, with percentages computed in Python."
)

cat(sprintf(
  "%.1f%% of candidate families appear in at least two recording windows.",
  py$repeated_percentage
))
```

## Interpretation and limitations

Recurrence across recording windows provides one way to prioritize
candidate frequency families for further review.

These percentages describe families, not buildings. Buildings contributing
more families have more influence on the summary, and windows from the
same recording are not independent building observations.

Recurrence alone does not confirm a structural mode. Directional evidence,
mode shapes, and other diagnostics remain separate review tasks.

The computational contribution here is a reproducible exchange:
R produces a summary, Python calculates from that summary, and R uses
the returned Python results in the same document.

## Acknowledgements

My broader doctoral research involves collaboration with
Dr. Carlos E. Ventura, Dr. Alemdar Bayraktar, and
Dr. Mehrtash Motamedi. This demonstration post presents a preliminary
sample of that research workflow.